Online Balanced Graph Partitioning: Optimizations for Large Clusters

### 2020-05-01 paper/disc20clustering.tex
Pass over intro

We consider two flavors of the problem
in this paper. In the first one, which
was also studied by Henzinger et al.~\cite{sigmetrics19_partitioning}
(SIGMETRICS 2019) and which
we here mainly consider for the sake of a stronger
lower bound, we assume that $\sigma$
simply reveals the edges of a static graph
that can be perfectly partitioned
(i.e., in principle, no external requests
are required). 
In the second one, we consider a general
model where $\sigma$ can be arbitrary;
this was the original problem introduced
by Avin et al.~\cite{repartition-disc} at DISC 2016.
For simplicity, we will refer to the former
model as the \emph{learning} model (as
one has to learn the static communication graph) 
and to the latter as the \emph{online} model.

\noindent \textbf{Contributions.}
We provide an improved lower bound 
of $\Omega(k\cdot\ell)$ on the competitive ratio of any online deterministic online algorithm 
even in the learning model;
the best known lower bound so far was $\Omega(k)$,
even for a more general model~\cite{repartition-disc}.
We also present an asymptotically optimal, 
$O(k\cdot \ell)$-competitive algorithm
for the restricted model.
For the online model, we present  
an asymptotically optimal,
$\Theta(\ell)$ algorithm for $k=3$;
the best known upper bound 
so far was $O(\ell^2)$~\cite{repartition-disc}.
%
Our algorithms can be distributed
similarly to the approach in~\cite{sigmetrics19_partitioning}.
The overview of known results is in Table~\ref{tab:overview}.

\maciek{TODO: fill the table: results and references to sections.}

@@

\noindent \textbf{Related work.}
The existing results in~\cite{repartition-disc}
and~\cite{sigmetrics19_partitioning}
primarily focus on a model where 
the online algorithm can use augmentation,
i.e., has slightly larger clusters than the offline 
algorithm. In contrast, in our paper, we focus
on a scenario where the nodes need to be
perfectly balanced among the clusters.
The problem has also been studied in a weaker
model where the adversary can only sample
requests from a fixed distribution~\cite{stochastic-ring}.
For clusters of size $2$ (i.e., $k=2$), 
it is known that a constant competitive algorithm exists~\cite{repartition-disc}.

@@

\noindent \textbf{Preliminaries.}

@@

>>>>

In addition to the general model, we study a simplified Perfect Partition model (also studied by Henzinger et al.~\cite{sigmetrics19_partitioning} (SIGMETRICS 2019)) that we consider for a stronger lower bound.
There, we assume that $\sigma$
reveals the edges of a static graph
that can be perfectly partitioned
(no external requests are required).

The general model, where $\sigma$ can be arbitrary
was the original problem introduced
by Avin et al.~\cite{repartition-disc} at DISC 2016.
For simplicity, we will refer to the Perfect Partition
model as the \emph{learning} model (as
one has to learn the static communication graph) 
and to the original as the \emph{repartitioning} model (as one continuously adjusts the partition in response to changing communication graph).

\noindent \textbf{Contributions.}
We provide an improved lower bound 
of $\Omega(k\cdot\ell)$ on the competitive ratio of any online deterministic online algorithm 
even in the learning model;
the best known lower bound so far was $\Omega(k)$,
even for a more general model~\cite{repartition-disc}.
We also present an asymptotically optimal, 
$O(k\cdot \ell)$-competitive algorithm
for the learning model.
For the repartitioning model, we present  
an asymptotically optimal,
$\Theta(\ell)$ algorithm for $k=3$;
the best known upper bound 
so far was $O(\ell^2)$~\cite{repartition-disc}.
Then, we present a strictly $6$-competitive algorithm for $k=2$.
In addition, we show a new, improved analysis of the algorithm from \cite{repartition-disc} that depends on the solutions to an interesting combinatorial problem (see Section~\ref{ssec:cascade}, The Cascade Problem).
%
The overview of known results is in Table~\ref{tab:overview}.

@@

\noindent \textbf{Related work.}
The existing results in~\cite{repartition-disc}
and~\cite{sigmetrics19_partitioning}
primarily focus on a model where 
the online algorithm can use augmentation,
i.e., has slightly larger clusters than the offline 
algorithm. In contrast, in our paper, we focus
on a scenario where the nodes need to be
perfectly balanced among the clusters.
The problem has also been studied in a weaker
model where the adversary can only sample
requests from a fixed distribution~\cite{stochastic-ring}.
For clusters of size $2$ (i.e., $k=2$), 
it is known that a $7$-competitive algorithm exists~\cite{repartition-disc} with large ($\Omega(\ell^2)$) additive constant.

@@

\section{Preliminaries.}

@@

Our algorithms can be distributed
similarly to the approach in~\cite{sigmetrics19_partitioning}.
